---
title: "Filter after normalizing"
author: "Divya Prima Crasta-237879"
date: "2024-11-12"
output: html_document
---

```{r}
library(mlr3)
library(mlr3extralearners)
library(mlr3measures)
library(mlr)
library(RWeka)
library(FSelector)
library(rpart)
library(ranger)
library(ggplot2)
library(dplyr)
library(tidyr)
```


```{r}
data_m2 <- read.csv("E:/Thesis/data/data_imputed_miss2.csv")
X <- as.matrix(data_m2[,-1])
# X is logarithmized, normalized, imputed(miss II)
y <- data_m2$Toxicity
```

# Assumption 1: Normality

for (i in 1:ncol(X)) {
  
  # Get the column name
  col_name <- colnames(X)[i]
  
  # Extract the mean and median for the current column from summary_df
  mean_val <- summary_df[col_name, "mean"]
  median_val <- summary_df[col_name, "median"]
  
  # Create the histogram and add vertical lines for mean and median
  gg <- ggplot(data_n_l, aes(x = get(col_name), fill = Toxicity)) + 
    geom_histogram( color = "black", binwidth = 1) + 
    xlab(col_name) + 
    theme_minimal() +
    labs(title = paste("Histogram of", col_name))
  
  # Print the plot
  print(gg)
}
```

```{r fig.width=20}
# Assumption2: Variance within each group is equal

# Convert to long format
long_data <- as.data.frame(X) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")
# Ensure 'Variable' is a factor and ordered based on its appearance in the dataset
long_data$Variable <- factor(long_data$Variable, levels = colnames(X))
long_data$Toxicity <- rep(y, each = ncol(X))

# Create boxplots for all numeric columns in one chart
ggplot(long_data, aes(x = Variable, y = Value, fill = Toxicity)) +
  geom_boxplot(outlier.color = "red") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Boxplots", x = "Variables", "Values")

## does not satisfy the variance homogeniety assumption
```

```{r}
# Convert to long format
long_data <- as.data.frame(X) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Ensure 'Variable' is a factor and ordered based on its appearance in the dataset
long_data$Variable <- factor(long_data$Variable, levels = colnames(X))

# Assign toxicity labels (assuming y is a vector of 0s and 1s)
long_data$Toxicity <- rep(y, each = ncol(X))
long_data$Toxicity <- as.factor(long_data$Toxicity)  # Ensure categorical

# Create boxplots for each variable with class 0 (blue) and class 1 (red)
ggplot(long_data, aes(x = Variable, y = Value, fill = Toxicity)) +
  geom_boxplot(outlier.color = "black", position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("0" = "blue", "1" = "red")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Boxplots by Variable and Toxicity Class", x = "Variables", y = "Values", fill = "Toxicity Class")
```

### Kruskal Test

```{r}
# Set the significance level
alpha <- 0.05

# Perform Kruskal-Wallis test for each variable and collect the results
kruskal_results <- apply(X, 2, function(var) {
  # Perform the test
  test_result <- kruskal.test(var ~ y)
  
  # Return 1 if p-value is less than alpha (reject null hypothesis), else 0
  ifelse(test_result$p.value < alpha, 1, 0)
})

# Convert results to a data frame with the desired format
kruskal_df <- as.data.frame(t(kruskal_results))
rownames(kruskal_df) <- "Kruskal"
colnames(kruskal_df) <- colnames(X)

# Display the results
print(kruskal_df)
```

### ROC-AUC

```{r}
aucs <- apply(X, 2, function(var) abs(0.5 - measureAUC(var, y, positive = 1, negative = 0)))
print(aucs)
print(sort(aucs, decreasing = TRUE))
```

### Feature Importance

```{r}
# Train a random forest model with impurity-based feature importance
rf_model <- ranger(
  formula = Toxicity ~ ., 
  data = data_n_l,  
  importance = "impurity",  # Get impurity-based importance
  classification = TRUE
)

# Extract impurity-based feature importance
impurity_importance <- rf_model$variable.importance

print(sort(impurity_importance, decreasing = TRUE))
```

### Final Result

```{r}
variables <- colnames(X)
# Create individual data frames, sorted by each metric
df_aucs <- data.frame(variable = variables, aucs = round(aucs, 4)) %>%
  arrange(desc(aucs))

# Combine the sorted data frames into one table
resultN <- data.frame(
  v_aucs = df_aucs$variable, aucs = df_aucs$aucs
)

# View the result
print(resultN)
```

```{r}
flags <- data.frame(matrix(0, nrow = nrow(resultN), ncol = ncol(resultN)%/%2))
rownames(flags) = colnames(X)
colnames(flags) = colnames(resultN)[seq(1, ncol(resultN), 2)]
for (col in colnames(flags)){
  for (variable in colnames(X)){
   if(variable %in% resultN[24,col]){
     flags[variable, col] = 1
   } 
  }
}

flags$sum <- rowSums(flags)

View(flags[order(flags$sum), c("sum"), drop = FALSE])


```

